### 2021-12-19 response_new.tex

\noindent Dear Editor:

We are re-submitting our manuscript, “The importance of the generation interval in investigating dynamics and control of new SARS-CoV-2 variants”, for consideration for publication in JRSI. This material has not previously been published elsewhere.

We had previously submitted the manuscript to JRSI (rsif-2021-0438) and were asked to revise with positive comments---for example, reviewer 3 said “The analysis is mathematically sound and offers a thorough investigation of an important concept that has been the subject of much discussion in the epidemiological community but, to my knowledge, hasn't been given the formal treatment it deserves until now.” We had submitted the revised version of the manuscript but the paper was rejected despite having addressed all comments due to reviewer 2's comment that “On a second reading of this manuscript, I am much less optimistic about its value to infectious disease epidemiology.” We emailed the editor since then and were provided the opportunity to resubmit.  
We have made revisions to our manuscript to address the reviewer's comments and to better highlight the novelty of our work.

The ongoing COVID-19 pandemic underlines the importance of characterizing the spread of new variants. Previous studies have particularly prioritized estimating the relative strength (i.e., the ratio of reproduction numbers) from relative speed (i.e., the differences between growth rates).
In doing so, analyses of new variants have focused on estimating the relative speed from the observed changes in the proportion of new variants while assuming that relative speed does not change over the course of invasion.
Here, we challenge this assumption by showing how relative strength and speed can change in response to different kinds of epidemic interventions and argue that holding relative strength fixed is generally more appropriate in the context of modeling the spread of SARS-CoV-2 variants.
In doing so, we consider the possibility that new variants can have different generation intervals from the original variant and explore how such differences affect the dynamics and control of new variants. 
We also show that neglecting these differences can lead to biased inference and provide a practical guidance for how such biases can be assessed in practice.
Below please find our detailed responses to reviewers.

@@

Sang Woo Park

\rev{Reviewer \#2}

\revtext{On a second reading of this manuscript, I am much less optimistic about its value to infectious disease epidemiology. The analysis is sound, and the revision does a better job of articulating the assumptions that are made along the way---addressing one of the main concerns in my previous review. However, the paper does not seem to make any particularly useful novel contribution to the field. The importance of assumptions about the generation interval distribution for estimates of reproduction numbers (based on growth rates) and growth rates (based on reproduction numbers) is well-established. The authors are right that this is often neglected in practice, and it is interesting to see how this plays out in their simulation studies.}

Thank you for your review.
We agree that the role of generation-interval distribution for estimates of reproduction numbers is well-established. We also agree that this is often neglected in practice. 
While we emphasize the role of generation-interval distribution in estimating transmission advantage of new variants, this conclusion is secondary to the main point of our paper. 
The goal is to challenge the idea of holding relative speed (differences in growth rates) constant.
In particular, we show that the relative speed changes under a constant-strength intervention, which reduces transmission rate by a constant amount and therefore keeps the relative strength constant---this effect is present even when the generation-interval distributions are identical and is exaggerated in the presence of differences in the generation-interval distributions.
We believe that using the generation-interval-based framework allows us to make this argument clearly.
We have re-written segments of the abstract, introduction, and discussion to make the contribution of our work clear.
We have also included brief reviews of more recent papers.

### 2021-12-19 newvariant.tex
changes changes changes

We have primarily focused on the differeces in the mean generation interval but the differences in the amount of variability, characeterized by the squared coefficient of variation $\kappa$, can also have important, but different implications \citep{blanquart2021selection}.
For example, if two variants have identical reproduction numbers, the variant with a wider generation-interval distribution will always have a faster growth rate and outcompete the other variant (except when both variants have $\RR=1$).
This is because a variant with a wider generation-interval distribution can take advantage of more early transmission during the growth phase and more late transmission during the decay phase.s

>>>>

We have primarily focused on the differences in the mean generation interval but the differences in the amount of variability, characterized by the squared coefficient of variation $\kappa$, can also have important, but different implications \citep{blanquart2021selection}.
For example, if two variants have identical reproduction numbers, the variant with a wider generation-interval distribution will always have a faster growth rate and out-compete the other variant (except when both variants have $\RR=1$).
This is because a variant with a wider generation-interval distribution can take advantage of more early transmission during the growth phase and more late transmission during the decay phase.s
